---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# readmedown

<!-- badges: start -->
<!-- badges: end -->

The goal of `readmedown` is render README files from Rmarkdown, with a particular focus on README files for data repositories.

## Installation

You can install the development version of `readmedown` like so:

``` r
# install.packages("remotes")
remotes::install_github("joelnitta/readmedown")
```

## Motivation

Data README files are very important sources of metadata that describe how other people can use the data you have made available in a repository (such as [FigShare](https://figshare.com/) or [Dryad](https://datadryad.org/stash)).

[Best practices for data README files](https://data.research.cornell.edu/content/readme) are to include as much information as possible about the format of each data file. This can be tedious to track manually. By using [Rmarkdown](https://bookdown.org/yihui/rmarkdown/), one can insert summary statistics, names of variables, etc., directly into the output document. This ensures your README files are accurate and saves time if/when your data changes!

## Usage

For example, the `desc_data()` function helps describe dataframes. It throws an error if your description of the data is wrong, thereby ensuring that the README is accurate:

```{r desc-data-example}
library(readmedown)

# Here are some example data: a small version of the 'mtcars' dataset
cars_small <- mtcars[1:10, 1:3]

# Let's describe these data with user-provided metadata
desc_data(
    cars_small,
    data.frame(
      col = c("mpg", "cyl", "disp"),
      desc = c("Miles per gallon", "Number of cylinders", "Displacement")
    )
  )
```


By using the `results = "asis"` [chunk option](https://yihui.org/knitr/options/), we can print the results of `desc_data()` directly into the rendered output. For example, you could produce the section below in a README file:

mtcars_small: A dataframe including `r nrow(cars_small)` rows and `r ncol(cars_small)` columns.

```{r desc-data-example-print, results = "asis", echo = FALSE}
desc_data(
    cars_small,
    data.frame(
      col = c("mpg", "cyl", "disp"),
      desc = c("Miles per gallon", "Number of cylinders", "Displacement")
    )
  )
```

## Template README file

This package includes a [template Rmd README file](inst/extdata/ReadmeTemplate.Rmd) based on this [awesome template by the Cornell Research Data Management Service Group](https://data.research.cornell.edu/content/readme).

I recommend using this template Rmd as a starting point for writing your data README file. It can be rendered to plain text using `plain_document()` provided by this package.

```{r template-example}
readme_rmd <- system.file(
  "extdata", "ReadmeTemplate.Rmd", package = "readmedown", mustWork = TRUE)
rmarkdown::render(readme_rmd)
```

## License

- [MIT](LICENSE.md)
